---
title: "Portfolio Annual Data Evaluation"
author: "Paul Leiby"
date: "April 20, 2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## We can look at the range of portfolio data

Accesible from from BetterInvesting

```{r getLibraries}
library(here)
library(tidyverse)
library(readr)
library(tidyquant)
library(knitr)  # for kable
```

```{r}
datapath = "data/"
```



```{r pressure, echo=FALSE}
# read set of stock data downloaded from ???
tkrData = read_delim("BOIC_Portfolio_AnnualData_20180506.txt",
                     delim="\t", comment="#")
```

```{r}
tkrData

```

http://tools.betterinvesting.org/cr/crreport.aspx?ticker=aapl

```{r}
tkrData %>% ggplot(aes(x=Year, y=HighPrice/EPS, colour=TKR)) + 
  geom_line() + ggtitle("Annual High PE Ratios")
```

### Gather stock price histories from Yahoo Finance

```{r readRecentValuation}
Valuation20200208 <- tribble(
~Security           , ~symbol,~FirstPurchase, ~SharesOwned, ~CostPerShr, ~TotalCost, ~CurrPrice, ~MarketValue, ~PctOfTotal,
"Amazon.com Inc       ","AMZN", "10/14/19", 10.0  ,1728.9100,17289.10 ,2079.28 ,20792.80 ,9.6,
"BJ Restaurants       ","BJRI", "04/15/19", 212.0 ,46.5228  ,9862.83  ,41.87   ,8876.44  ,4.1,
"Columbia Sportswear  ","COLM", "11/11/19", 150.0 ,93.3900  ,14008.50 ,90.45   ,13567.50 ,6.3,
"Canadian Pacific Rai ","CP",   "06/20/16", 60.0  ,148.4933 ,8909.60  ,267.00  ,16020.00 ,7.4,
"G III Apparel Group  ","GIII", "03/04/19", 332.0 ,36.0709  ,11975.55 ,27.76   ,9216.32  ,4.3,
"Google Alphabet Clas ","GOOG", "08/11/08", 16.0  ,931.2044 ,14899.27 ,1479.23 ,23667.68 ,11.0,
"Graphc Pack Hldg     ","GPK",  "09/10/18", 424.0 ,14.1664  ,6006.55  ,15.86   ,6724.64  ,3.1,
"Healthcare Services  ","HCSG", "03/12/18", 200.0 ,47.4448  ,9488.95  ,26.42   ,5284.00  ,2.5,
"Jazz Pharmaceuticals ","JAZZ", "11/21/16", 76.0  ,131.0114 ,9956.87  ,142.54  ,10833.04 ,5.0,
"LKQ                  ","LKQ",  "09/15/14", 658.0 ,27.3488  ,17995.52 ,32.93   ,21667.94 ,10.1,
"Mastec Inc           ","MTZ",  "01/16/18", 90.0  ,53.2272  ,4790.45  ,56.86   ,5117.40  ,2.4,
"Nvidia Corp          ","NVDA", "02/11/19", 68.0  ,146.4922 ,9961.47  ,251.59  ,17108.12 ,7.9,
"Skyworks Solutions   ","SWKS", "02/10/14", 104.0 ,96.3368  ,10019.03 ,113.94  ,11849.76 ,5.5,
"Thor Industries Inc  ","THO",  "08/13/18", 102.0 ,96.4381  ,9836.69  ,81.23   ,8285.46  ,3.8,
"Totals               ","total", NA       , NA      , NA      ,155000.38, NA,179011.10 ,83.1
)

Valuation20200419 <- tribble(
  ~Security           , ~symbol,~FirstPurchase, ~SharesOwned, ~CostPerShr, ~TotalCost, ~CurrPrice, ~MarketValue, ~PctOfTotal,
"BJ Restaurants       ","BJRI", "04/15/19" ,212.0 , 46.5228 , 9862.83  ,17.39  , 3686.68 ,2.0,
"Columbia Sportswear  ","COLM", "11/11/19" ,150.0 , 93.3900 , 14008.50 ,72.44  , 10866.00,6.0,
"G III Apparel Group  ","GIII", "03/04/19" ,332.0 , 36.0709 , 11975.55 ,9.34   , 3100.88 ,1.7,
"Graphc Pack Hldg     ","GPK",  "09/10/18" ,424.0 , 14.1664 , 6006.55  ,13.18  , 5588.32 ,3.1,
"Healthcare Services  ","HCSG", "03/12/18" ,200.0 , 47.4448 , 9488.95  ,20.92  , 4184.00 ,2.3,
"Jazz Pharmaceuticals ","JAZZ", "11/21/16" ,76.0  , 131.0114, 9956.87  ,111.19 , 8450.44 ,4.7,
"LKQ                  ","LKQ",  "09/15/14" ,565.0 , 26.3850 , 14907.53 ,20.94  , 11831.10,6.6,
"Mastec Inc           ","MTZ",  "01/16/18" ,90.0  , 53.2272 , 4790.45  ,34.80  , 3132.00 ,1.7,
"Skyworks Solutions   ","SWKS", "02/10/14" ,104.0 , 96.3368 , 10019.03 ,95.40  , 9921.60 ,5.5,
"Thor Industries Inc  ","THO",  "08/13/18" ,102.0 , 96.4381 , 9836.69  ,53.27  , 5433.54 ,3.0,
"Totals               ","total", NA        , NA      , NA       ,100852.95 , NA, 66194.56 ,36.8
)
```


```{r}
Valuation <- Valuation20200419 %>%
  filter(symbol != "total")
ValuationDate = as.Date(2020/04/19)

BOICtickers <- Valuation20200208 %>% # use current or prior portfolio
  filter(symbol != "total") %>%
  select(symbol, Security) 
# BOICtickers <- tribble(
#   ~ symbol,  ~ full_names,
#   "AMZN",    "Amazon",
#   "AAPL",    "Apple",
#   "BJRI",    "BJ Restaurants",
#   "COLM",    "Columbia Sportswear",
#   "CP",      "Canadian Pacific",
#   "CRI",     "Carters Inc",
#   "DISCA",   "Discovery Comm",  
#   "DKS",     "Dicks Sporting Goods",
#   "DY",      "Dycom", 
#   "GIL",     "Gilden",
#   "GOOG",    "Google",
#   "JAZZ",    "Jazz Pharma",
#   "LKQ",     "LKQ",
#   "PHM",     "Pulte",
#   "QCOM",    "Qualcomm",
#   "R",       "Ryder")
```

```{r displayPortfolio}
Valuation %>%
  kable(caption = paste0("Recent Portfolio Valuation, as of ", ValuationDate))
```

```{r}
BOICtickersHistorical <- tribble(
  ~ symbol,  ~ full_names,
  "AAPL",    "Apple",
  "CBI",     "xxx",
  "CMI",     "xxx",
  "EMC",     "xxx",
  "GIL",     "xxx",
  "HAR",     "xxx",
  "LKQ",     "xxx",
  "MSM",     "xxx",
  "QCOM",    "xxx",
  "R",       "xxx",
  "WFM",     "xxx",
  "SNCR",    "xxx",
  "PWR",     "xxx",
  "SWKS",    "xxx",
  "UNP",     "xxx",
  "GOOG",    "xxx",
  "GOOGL",   "xxx"
)

IndexTickers <- tribble(
  ~ symbol,  ~ full_names,
  "^GSPC"	,  "S&P500",
  "^DJI"	,  "Dow_30",
  "^IXIC" ,  "Nasdaq",
  "^RUT"	,  "Russell2000",
  "^VIX"	,  "Vix",
  "^GDAXI",  "DAX_index",
  "^HSI"  ,  "HangSengIndex"
)

```

Now we pass the tickers to tidyquant::tq_get().

```{r}
refreshPriceData = F # plan to read price data from the web

if (refreshPriceData) {
  tkr_price_data <- 
    tq_get(BOICtickers$symbol,                         
           get = "stock.prices",             
           from = "1994-01-01",
           complete_cases = F)
  
  indx_price_data <- 
    tq_get(IndexTickers$symbol,                         
           get = "stock.prices",             
           from = "1994-01-01",
           complete_cases = F)
} else {
  tkr_price_data <- read_csv(file = here::here(datapath,"BOICtickersPrices.csv"))
  indx_price_data <- read_csv(file = here::here(datapath,"indxtickersPrices.csv"))
}


```

We have our data and group by the symbol column, displaying the top row for each symbol.

```{r}
tkr_price_data %>% 
  group_by(symbol) %>% 
  slice(1) # show the first record of each group
```

Save the downloaded data

```{r writeTkrDataToCSV, include=F}
if (refreshPriceData) { # save newly-read data
  tkr_price_data %>% select(date, symbol, adjusted) %>%
    write.csv(file = here::here(datapath,"BOICtickersPrices.csv"))
  indx_price_data %>% select(date, symbol, adjusted) %>%
    write.csv(file = here::here(datapath,"indxtickersPrices.csv"))
}
```

We want to visualize the period-to-period change in prices.

We are not doing any annualizing or seasonality work here – it’s a simple substraction.
For yearly change, it would be value - lag(value, NunPeriodsPerYear).

```{r compute }
all_price_data <- tkr_price_data %>%
  bind_rows(indx_price_data)

tkr_daily_change <- 
  all_price_data  %>% 
  select(symbol, date, adjusted) %>%
  group_by(symbol) %>%  # grouping needed to not lag across adjacent observations for 2 groups
  mutate(price_change = log(adjusted) - lag(log(adjusted), 1)) %>% 
  na.omit()
```


#### Let’s build some charts.

We will start at the top and use ggplot.
We want positive growth periods to be one color, and negative to be another. 

Next, we create two new columns with mutate(). The first is called col_pos and is formed by if_else(price_change > 0, price_change,...). That logic is creating a column that holds the value of monthly change if monthly change is positive, else it holds NA. We then create another column called col_neg using the same logic.

```{r calcDailyPriceChange}
price_change = tkr_daily_change %>% 
   mutate(col_pos = 
           if_else(price_change > 0, 
                  price_change, as.numeric(NA)),
         col_neg = 
           if_else(price_change < 0, 
                  price_change, as.numeric(NA)))

price_change %>%
  dplyr::select(symbol, date, col_pos, col_neg) %>%
  filter(symbol == "AMZN")
```

Have a qucik look at the col_pos and col_neg columns and make sure they look right. col_pos should have only positive and NA values, col_neg shoud have only negative and NA values.



An alternative is to use highcharter to create an interactive chart, responding to mouse-overs.

```{r calcMonthlyPriceChange}
mprice_change <- 
  all_price_data  %>% 
  select(symbol, date, adjusted) %>%
  mutate(lnprice = log(adjusted)) %>%
  mutate(month = month(date),
         year = year(date)) %>%
  group_by(symbol) %>%  # grouping needed to not lag across adjacent observations for 2 groups
  mutate(dlnprice = lnprice - lag(lnprice, 1)) %>% 
  na.omit() %>%
  group_by(symbol, year, month) %>%
  summarize(m_dln_price = sum(dlnprice)) %>%
  ungroup() %>%
  # may be a better function in lubridate for this date conversion
  mutate(date = as.Date(paste(year, month, "15", sep='-')))
  
mprice_change = mprice_change %>% 
   mutate(
     col_pos = 
       if_else(m_dln_price > 0, m_dln_price, as.numeric(NA)),
     col_neg = 
       if_else(m_dln_price < 0, 
               m_dln_price, as.numeric(NA)))

```

Display Monthly Price Changes Graphically

```{r}
plotUpsAndDown  <- function(df, 
                            currsymbol = "", 
                            currtitle = "Fractional Price Change", 
                            currsubtitle = "Monthly Prices",
                            earliestDataWanted = "1960/01/01") {
  earliest = as.Date(earliestDataWanted)
  df %>% 
    filter(date >= earliest) %>%
    filter(symbol == currsymbol) %>%
    select(symbol, date, col_pos, col_neg) %>%
    ggplot(aes(x = date)) +
    geom_col(aes(y = col_neg),
                 alpha = .85,
                 fill = "pink",
                 color = "pink") +
    geom_col(aes(y = col_pos),
                 alpha = .85,
                 fill = "lightgreen",
                 color = "lightgreen") +
    ylab("Change (Pct/month)") +
    labs(title = currtitle,
         subtitle = paste0(currsubtitle, ", ", currsymbol),
         caption = "Fractional price changes (in $/share)%/100") +
    scale_x_date(breaks = scales::pretty_breaks(n = 10)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5),
          plot.caption = element_text(hjust=0))
}
```

```{r}
plotUpsAndDown(df = price_change, currsymbol = "AMZN", currtitle = "Fractional Price Change", currsubtitle = "Daily Prices") 
```

```{r}
plotUpsAndDown(df = mprice_change, currsymbol = "AMZN", currtitle = "Fractional Price Change", currsubtitle = "Monthly Prices") 
```

```{r}
plotUpsAndDown(df = price_change, currsymbol = "BJRI", 
               currtitle = "Fractional Price Change", 
               currsubtitle = "Daily Prices") 
```

```{r}
plotUpsAndDown(df = mprice_change, currsymbol = "BJRI", 
               currtitle = "Fractional Price Change", 
               currsubtitle = "Monthly Prices",
               earliestDataWanted = "2020/01/01") 
```


```{r}
plotUpsAndDown(df = price_change, currsymbol = "BJRI", 
               currtitle = "Fractional Price Change", 
               currsubtitle = "Daily Prices",
               earliestDataWanted = "2020/01/01") 
```
### Compare price movements acroos securities

```{r defFn_plotRelativePrices}
plotRelativePrices  <- function(df, 
                            currSymbolsWanted = c("^GSPC"),
                            currtitle = "Relative Prices", 
                            currsubtitle = "Monthly Prices",
                            earliestDateWanted = "1960/01/01",
                            relativeP = F) {

  earliest <- as.Date(earliestDateWanted)
  df <- df %>%  # restrict to dates and tickers wanted
    filter(date >= earliest) %>%
    filter(symbol %in% currSymbolsWanted)

  earliest <- df %>% 
    # ungroup() %>%
    select(date) %>%
    pull() %>%
    min() # this makes sure the the earliest actually a trade date with obs

  if (relativeP == T) {
    df <- df %>%
    group_by(symbol) %>%
    mutate( # calculate relative prices compared to earliest date wanted
      adjusted = adjusted/adjusted[date == earliest]
    )
  }

  df %>% # plot this up
    group_by(symbol) %>%
    ggplot(aes(x = date, y=adjusted, color=symbol)) +
    ylab("Relative Price") +
    labs(title = currtitle,
         subtitle = paste0(currsubtitle),
         caption = paste0("Prices over starting Price on ", earliest)) +
    geom_line() +
    scale_x_date(breaks = scales::pretty_breaks(n = 10)) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1),
          plot.title = element_text(hjust = 0.5),
          plot.subtitle = element_text(hjust = 0.5),
          plot.caption = element_text(hjust=0))
}
```

```{r}
stocksWanted = c("^GSPC", "AMZN", "NVDA", "GOOG", "^VIX")

plotRelativePrices(price_change, currSymbolsWanted = stocksWanted, 
                 earliestDateWanted = "2019/01/01",  relativeP = T)
```

```{r}
plotRelativePrices(price_change, currSymbolsWanted = stocksWanted, 
                 earliestDateWanted = "2020/01/01",  relativeP = T)
```

```{r}
stocksWanted = c("^GSPC", BOICtickers$symbol)
stocksWanted

plotRelativePrices(price_change, currSymbolsWanted = stocksWanted, 
                   earliestDateWanted = "2020/01/03",  relativeP = T)
```

