---
title: "Portfolio Annual Data Evaluation"
author: "Paul Leiby"
date: "Aug 11, 2018"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## We can look at the range of portfolio data

Accesible from from BetterInvesting

```{r cars}
library(tidyverse)
library(readr)
library(tidyquant)
```


```{r pressure, echo=FALSE}
# read set of stock data downloaded from ???
tkrData = read_delim("BOIC_Portfolio_AnnualData_20180506.txt",
                     delim="\t", comment="#")
```

```{r}
tkrData

```

```{r}
tkrData %>% ggplot(aes(x=Year, y=HighPrice/EPS, colour=TKR)) + 
  geom_line() + ggtitle("Annual High PE Ratios")
```

### Gather stock price histories from Yahoo Finance

```{r}
BOICtickers <- tribble(
  ~ symbol,  ~ full_names,
  "AAPL",    "Apple",
  "CMI",     "Cummins",
  "CP",      "Canadian Pacific",
  "CRI",     "Carters Inc",
  "DISCA",   "Discovery Comm",  
  "DKS",     "Dicks Sporting Goods",
  "DY",      "Dycom", 
  "GIL",     "Gilden",
  "GOOG",    "Google",
  "JAZZ",    "Jazz Pharma",
  "LKQ",     "LKQ",
  "PHM",     "Pulte",
  "QCOM",    "Qualcomm",
  "R",       "Ryder")

BOICtickersHistorical <- tribble(
  ~ symbol,  ~ full_names,
  "AAPL",    "xxx",
  "CBI",     "xxx",
  "CMI",     "xxx",
  "EMC",     "xxx",
  "GIL",     "xxx",
  "HAR",     "xxx",
  "LKQ",     "xxx",
  "MSM",     "xxx",
  "QCOM",    "xxx",
  "R",       "xxx",
  "WFM",     "xxx",
  "SNCR",    "xxx",
  "PWR",     "xxx",
  "SWKS",    "xxx",
  "UNP",     "xxx",
  "GOOG",    "xxx",
  "GOOGL",   "xxx"
)

IndexTickers <- tribble(
  ~ symbol,  ~ full_names,
  "^GSPC"	,  "S&P500",
  "^DJI"	,  "Dow_30",
  "^IXIC" ,  "Nasdaq",
  "^RUT"	,  "Russell2000",
  "^VIX"	,  "Vix",
  "^GDAXI",  "DAX_index",
  "^HSI"  ,  "HangSengIndex"
)

```

Now we pass the tickers to tidyquant::tq_get().

```{r}
tkr_price_data <- 
  tq_get(BOICtickers$symbol,                         
         get = "stock.prices",             
         from = "1994-01-01",
         complete_cases = F)

indx_price_data <- 
  tq_get(IndexTickers$symbol,                         
         get = "stock.prices",             
         from = "1994-01-01",
         complete_cases = F)

```

We have our data and group by the symbol column, displaying the top row for each symbol.

```{r}
tkr_price_data %>% 
  group_by(symbol) %>% 
  slice(1)
```

We want to visualize the month-to-month change in jobs numbers, which means we need to perform a calculation on our data and store it in a new column. We use mutate() to create the new column and calculate monthly change with value - lag(value, 1). We are not doing any annualizing or seasonality work here – it’s a simple substraction. For yearly change, it would be value - lag(value, 12).

```{r}
tkr_daily_change <- 
  tkr_price_data  %>% 
  select(symbol, date, adjusted) %>%
  group_by(symbol) %>%  # grouping needed to not lag across adjacent observations for 2 groups
  mutate(price_change = adjusted - lag(adjusted, 1)) %>% 
  na.omit()
```


#### Let’s build some charts.

We will start at the top and use ggplot.
We want positive growth periods to be one color, and negative to be another. 

Next, we create two new columns with mutate(). The first is called col_pos and is formed by if_else(price_change > 0, price_change,...). That logic is creating a column that holds the value of monthly change if monthly change is positive, else it holds NA. We then create another column called col_neg using the same logic.

```{r}
price_change = tkr_daily_change %>% 
  filter(symbol == "AAPL") %>% 
   mutate(col_pos = 
           if_else(price_change > 0, 
                  price_change, as.numeric(NA)),
         col_neg = 
           if_else(price_change < 0, 
                  price_change, as.numeric(NA))) %>% 
  dplyr::select(symbol, date, col_pos, col_neg)

price_change
```

Have a qucik look at the col_pos and col_neg columns and make sure they look right. col_pos should have only positive and NA values, col_neg shoud have only negative and NA values.

Now we can visualize our monthly changes with ggplot, adding a separate geom for those new columns.

```{r}
price_change %>% 
  ggplot(aes(x = date)) +
  geom_col(aes(y = col_neg),
               alpha = .85,
               fill = "pink",
               color = "pink") +
  geom_col(aes(y = col_pos),
               alpha = .85,
               fill = "lightgreen",
               color = "lightgreen") +
  ylab("Change ($/share)") +
  labs(title = "Price Change",
       subtitle = "Daily Prices",
       caption = "Nominal price changes in $/share") +
  scale_x_date(breaks = scales::pretty_breaks(n = 10)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5),
        plot.caption = element_text(hjust=0))
```

An alternative is to use highcharter to create an interactive chart, responding to mouse-overs.

