---
title: "test tidyquant"
author: "Paul N. Leiby"
date: "4/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
# Load required packages.
library(tidyr)
library(dplyr)
library(tibble)
```

```{r}
# Custom R function as Data.
Stock3.func <- function(){
  require(quantmod)
  google <- rownames_to_column(data.frame(getSymbols("GOOG", auto.assign = FALSE)))
  facebook <- rownames_to_column(data.frame(getSymbols("FB", auto.assign = FALSE)))
  twitter <- rownames_to_column(data.frame(getSymbols("TWTR", auto.assign = FALSE)))
  amazon <- rownames_to_column(data.frame(getSymbols("AMZN", auto.assign = FALSE)))
  msft <- rownames_to_column(data.frame(getSymbols("MSFT", auto.assign = FALSE)))
  apple <- rownames_to_column(data.frame(getSymbols("AAPL", auto.assign = FALSE)))
  bind_rows(google, facebook,amazon, apple, twitter, msft)
}
# Data Analysis Steps
Stock3.func() %>%
  gather(measures, value, -rowname, na.rm = TRUE) %>%
  separate(measures, into = c("symbol", "measure"), sep = "\\.", remove = TRUE, convert = FALSE, extra = "drop", fill = "right") %>%
  spread(measure, value) %>%
  rename(date = rowname)
```

`tq_get`

- "economic.data": Get economic data from FRED. rapper for quantmod::getSymbols.FRED()

- "quandl": Get data sets from Quandl. Wrapper for Quandl::Quandl(). See also quandl_api_key().

- "tiingo": Get data sets from Tiingo. Wrapper for riingo::riingo_prices(). See also tiingo_api_key().

- "tiingo.iex": Get data sets from Tiingo. Wrapper for riingo::riingo_iex_prices(). See also tiingo_api_key().

- "tiingo.crypto": Get data sets from Tiingo. Wrapper for riingo::riingo_crypto_prices(). See also tiingo_api_key().

- "alphavantager": Get data sets from Alpha Vantage. Wrapper for alphavantager::av_get(). See also av_api_key().

```{r}
library(tidyquant)

SP500composition <- tq_index(x = "SP500")

SP500prices <- tq_index(x = "SP500") %>%
  tq_get(get = "stock.prices")

SP500prices <- SP500prices %>%
  select(symbol, company, date, adjusted)

write_csv(x = SP500prices, here::here(datapath, "SP500prices.csv"))
write_csv(x = SP500composition, here::here(datapath, "SP500composition.csv"))
```

```{r}
# quantmod::getSymbols.FRED()
```

